Portfolios of Subgraph Isomorphism Algorithms

### 2015-11-30 paper/paper.tex
some background, experimental setup, results

@@

We could try to come up with rules to pick an algorithm on an instance by instance basis using our
knowledge of how the algorithms behave: for example, maybe we should use the Glasgow algorithm with
paths of length 2 with large target graphs (but how large?), and only use length 3 on smaller
graphs. Also, we could guess that LAD might do better than Glasgow on smaller, heavily-structured
graphs. However, this is not very systematic.

?? Lars: a paragraph of fluff about instance selection please.

We show that picking algorithms and parameter choices on an instance by instance basis, by
considering features of the pattern and target graphs, gives better overall performance than any
single algorithm. This is the first time this has been done for subgraph isomorphism, and we must
consider an unusual factor: our features must look at pairs of inputs, and the relationships between
them.

?? Lars: do you know of any other problems where you have two inputs rather than one? Or is this
novel?

@@

@@

?? Lars: your stuff here.

@@

\noindent Note that this is not like SAT, where we can reasonably split things up based upon which
family they're from.

\subsection{Machine learning voodoo stuff}

\subsection{Results}

?? Can we do better than picking the single best on a family by family basis? Bear in mind that we
don't have the family as a usable feature.

>>>>

\usepackage{todonotes}
\usepackage{booktabs}

@@

We show that by choosing algorithms on an instance by instance basis, by
considering features of the pattern and target graphs, we achieve better overall
performance than any single algorithm.

@@

\section{Background}

\todo[inline]{Some background on subgraph isomorphism algorithms, maybe move
some stuff from the intro here.}

The per-instance algorithm selection problem~\cite{rice_algorithm_1976} is to
select from an algorithm
portfolio~\cite{huberman_economics_1997,gomes_algorithm_2001} the one expected
to perform best on a given problem instance. Algorithm selection systems usually
build machine learning models of the algorithms or the portfolio they are
contained in to forecast which algorithm to use in a particular context. Using
the model predictions, one or more algorithms from the portfolio are selected to
be run sequentially or in parallel.

Here, we consider the case where exactly one algorithm is selected for
solving the problem. One of the most prominent and successful systems that
employs this approach is SATzilla~\cite{xu_satzilla_2008}, which defined the
state of the art in SAT solving for a number of years. Other application areas
include constraint solving~\cite{omahony_using_2008},
the travelling salesperson problem~\cite{kotthoff_improving_2015},
and AI planning~\cite{seipp_learning_2012}.

The interested reader is referred to a recent
survey~\cite{kotthoff_algorithm_2014} for additional information on algorithm
selection.

@@

@@

We consider a total of 5725 instances.

\subsubsection{Features}

\todo[inline]{Describe features here.}

This gives us a total of 30 features.

The data we use in this paper is available as Aslib scenario GRAPHS-2015.

\subsection{Algorithm Selection approach}

We apply two static steps to each instance before solving it. First, the
computation of the LAD features can detect unsatisfiability of an instance. If
an instance is identified as unsatisfiable, we do not process it further.
Second, we run the VF2 solver for 0.05 seconds. This solves easy instances
without the overhead of running algorithm selection and avoids potentially
making incorrect solver choices.

Computing the LAD features detects unsatisfiability on 1919 instances. Running
VF2 for 0.05 seconds removes a further 1470. We are left with 2336 hard
instances that we consider for algorithm selection.

To decide which solver to run on the remaining instances, we use
LLAMA~\cite{kotthoff_llama_2013}. In preliminary experiments, we identified
random forest regression to predict the performance difference between pairs of
algorithms, with the best algorithm identified as the one with the best
cumulative performance difference over all pairs, as the best approach. We make
no claims that this is the overall best approach.

\section{Results}

Table~\ref{tab:res} shows the performance of our algorithm selection approach,
compared to two baselines. The virtual best solver is the oracle predictor that,
for each instance, chooses the best solver from our portfolio. This is the upper
bound of what an algorithm selection approach can achieve. The single best
solver is the one solver from the portfolio that has the overall best
performance across the entire set of instances. We consider it a lower bound on
performance.

\begin{table}[ht]
\centering
\begin{tabular}{lr}
  \toprule
& mean runtime\\
  \midrule
virtual best & 5822.809\\
  single best & 7782.569\\
  LLAMA & 6595.275\\
   \bottomrule
\end{tabular}
\vspace{1ex}
\caption{Algorithm selection performance on reduced set of 2336
instances.}\label{tab:res}
\end{table}

We are able to close more than 60\% of the performance gap between the single
best and the virtual best solver, our lower and upper bounds. For comparison, we
show the results achieved on the entire set of instances without
unsatisfiability detection and presolving in Table~\ref{tab:resfull}. We achieve
similar results with more than 58\% of the gap between the single best and the
virtual best closed.

\begin{table}[ht]
\centering
\begin{tabular}{lr}
  \toprule
& mean runtime\\
  \midrule
virtual best & 2375.913\\
  single best & 3177.569\\
  LLAMA & 2706.330\\
   \bottomrule
\end{tabular}
\vspace{1ex}
\caption{Algorithm selection performance on the full set of 5725
instances.}\label{tab:resfull}
\end{table}

Figure~\ref{fig:scatter} shows detailed results for each of the 2336 instances
in the reduced set. Our algorithm selection approach is by no means able to
identify the best solver on a majority of instances, although we are able to
achieve performance improvements overall. There seems to be no particular
pattern as to where algorithm selection performs well and where it fails.

\begin{figure}[!ht]
\includegraphics[width=\textwidth]{figures/perfScatter}
\caption{Single best algorithm vs.\ portfolio performance on all instances by
family. Points above the diagonal indicate that the portfolio achieved better
performance, while the single best solver was better on instances below the
diagonal.}
\label{fig:scatter}
\end{figure}
